---
title: "Scripts for: Invasive swamp eels do not elicit antipredator behavior in a native crayfish with collapsing populations "
author: "Brandon Güell"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 6
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 6
---

```{r setup, include=FALSE}
#setup PDF Knit settings for export as PDF [this entire chunk will not show on knit doc]
knitr::opts_chunk$set(echo = TRUE, max.print=10, message = FALSE, warning = FALSE, tidy = T, fig.show = "hold") # display r chunks but not messages (e.g., warnings) in knit doc

# Load required packages
library(readxl)
library(dplyr)
library(zoo)
library(ggplot2)
library(scales)
library(lemon)
library(lme4)
library(mgcv)
library(car)
library(lmerTest)
library(emmeans)
library(glmmTMB)
library(DHARMa)
library(epitools)
library(epiR)
library(gridExtra)
library(cowplot)
```

# Introduction

> The data processing and statistical analyses shown here are from the Guell & Dorn "Invasive swamp eels do not elicit antipredator behavior in a native crayfish with collapsing populations" manuscript submitted to Proceedings B (manuscript ID RSPB-2025-0356). Analyses are shows in chronological order as reported in the manuscript. We show details on importing, manipulating, summarizing, analyzing (testing hypotheses), and visualizing the data. See manuscript for details on background, research questions, hypothesis, methods, etc..


# Nonlethal Predator Experiment

Import dataset
```{r}
library(readxl)
cray <- read_excel("Datasets for Invasive swamp eels do not elicit antipredator behavior in a native crayfish with collapsing populations.xlsx", 
    sheet = "Nonlethal Predator Experiment", 
    na = "NA")

craycray = cray %>%
  mutate(Treatment_Period = factor(Treatment_Period, levels=c("Before", "After")))
```


## Crayfish morphometrics

> Rough estimates of CL from crayfish parts in eel diets (N=9) sizes were 8.9±3.2 mm (mean±SD) and ranged from 3.7–14.5 mm. We tried to test crayfish that were below 15 mm in CL.

### Summary statistics
```{r}
cray_size_summary0 = cray %>% 
summarise(mean = mean(Crayfish_CL_mm, na.rm=T), SE= sd(Crayfish_CL_mm, na.rm=T)/sqrt(length(Crayfish_CL_mm)), SD = sd(Crayfish_CL_mm, na.rm=T), max = max(Crayfish_CL_mm, na.rm=T), min = min(Crayfish_CL_mm, na.rm=T), sample_size_unique=n_distinct(Crayfish_CL_mm, na.rm=T), sample_size=n())
cray_size_summary0

cray_size_summary0.5 = craycray %>% 
  group_by(Predator_Treatment) %>%
summarise(mean = mean(Crayfish_CL_mm, na.rm=T), SE= sd(Crayfish_CL_mm, na.rm=T)/sqrt(length(Crayfish_CL_mm)), SD = sd(Crayfish_CL_mm, na.rm=T), max = max(Crayfish_CL_mm, na.rm=T), min = min(Crayfish_CL_mm, na.rm=T), sample_size_unique=n_distinct(Crayfish_CL_mm, na.rm=T), sample_size=n())
cray_size_summary0.5

cray_size_summary1 = craycray %>% 
  group_by(Predator_Treatment, Population) %>%
summarise(mean = mean(Crayfish_CL_mm, na.rm=T), SE= sd(Crayfish_CL_mm, na.rm=T)/sqrt(length(Crayfish_CL_mm)), SD = sd(Crayfish_CL_mm, na.rm=T), max = max(Crayfish_CL_mm, na.rm=T), min = min(Crayfish_CL_mm, na.rm=T), sample_size_unique=n_distinct(Crayfish_CL_mm, na.rm=T), sample_size=n())
cray_size_summary1

cray_size_summary2 = craycray %>% 
  group_by(Crayfish_Sex, Population, Predator_Treatment) %>%
summarise(mean = mean(Crayfish_CL_mm, na.rm=T), SE= sd(Crayfish_CL_mm, na.rm=T)/sqrt(length(Crayfish_CL_mm)), SD = sd(Crayfish_CL_mm, na.rm=T), max = max(Crayfish_CL_mm, na.rm=T), min = min(Crayfish_CL_mm, na.rm=T), sample_size_unique=n_distinct(Crayfish_CL_mm, na.rm=T), sample_size=n())
cray_size_summary2
```

> Crayfish tested were 8.7-17 mm in CL and were 12.43+/-1.54



*********************************************************************************************************
*********************************************************************************************************
*********************************************************************************************************


## TIME ACTIVE
```{r}
craycray_after_period_only = craycray %>%
  filter(Treatment_Period == "After")

time_active_craycray_summary1 = craycray %>% 
  group_by(Predator_Treatment, Treatment_Period) %>%
summarise(mean = mean(Time_Active_percent, na.rm=T), SE= sd(Time_Active_percent, na.rm=T)/sqrt(length(Time_Active_percent)), SD = sd(Time_Active_percent, na.rm=T), max = max(Time_Active_percent, na.rm=T), min = min(Time_Active_percent, na.rm=T), sample_size_unique=n_distinct(Time_Active_percent, na.rm=T), sample_size=n(), ci_lower = mean-qt(0.975, df = n() - 1) *SE, ci_upper = mean + qt(0.975, df = n() - 1) * SE)
time_active_craycray_summary1

time_active_craycray_summary2 = craycray %>% 
  group_by(Population, Predator_Treatment, Treatment_Period) %>%
summarise(mean = mean(Time_Active_percent, na.rm=T), SE= sd(Time_Active_percent, na.rm=T)/sqrt(length(Time_Active_percent)), SD = sd(Time_Active_percent, na.rm=T), max = max(Time_Active_percent, na.rm=T), min = min(Time_Active_percent, na.rm=T), sample_size_unique=n_distinct(Time_Active_percent, na.rm=T), sample_size=n())
time_active_craycray_summary2

time_active_craycray_summary3 = craycray %>% 
  group_by(Treatment_Period) %>%
summarise(mean = mean(Time_Active_percent, na.rm=T), SE= sd(Time_Active_percent, na.rm=T)/sqrt(length(Time_Active_percent)), SD = sd(Time_Active_percent, na.rm=T), max = max(Time_Active_percent, na.rm=T), min = min(Time_Active_percent, na.rm=T), sample_size_unique=n_distinct(Time_Active_percent, na.rm=T), sample_size=n())
time_active_craycray_summary3

time_active_craycray_summary4 = craycray %>% 
  group_by(Population, Predator_Treatment, Treatment_Period, Crayfish_Sex) %>%
summarise(mean = mean(Time_Active_percent, na.rm=T), SE= sd(Time_Active_percent, na.rm=T)/sqrt(length(Time_Active_percent)), SD = sd(Time_Active_percent, na.rm=T), max = max(Time_Active_percent, na.rm=T), min = min(Time_Active_percent, na.rm=T), sample_size_unique=n_distinct(Time_Active_percent, na.rm=T), sample_size=n())
time_active_craycray_summary4
```

#### Visualizations
```{r}
figure1 = 
  ggplot(time_active_craycray_summary1, aes(x=Treatment_Period, y=mean)) +
  geom_path(aes(color=Predator_Treatment, group = Predator_Treatment), position = position_dodge(width=0.1), size = 3) +
  geom_errorbar(aes(color=Predator_Treatment, ymin=mean-SE, ymax=mean+SE), position = position_dodge(width=0.1), width=0.5, size=2) +
  geom_point(aes(fill=Predator_Treatment, shape = Predator_Treatment), position = position_dodge(width=0.1), size = 3, color = "black") +
  #stat_summary(aes(color=Predator_Treatment, alpha=Population), fun=mean, geom='crossbar', position = position_jitterdodge(.1), size=1, linewidth=.5) +
  labs(y = "Mean time active (±95% CI)", x="Treatment") + #creat labels... 
  theme_classic(base_size=12) +
  theme(legend.text = element_text(size = 12), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 12), axis.ticks = element_line(color = "black"), axis.title.x = element_blank()) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= c(.05,.1), legend.title = element_blank()) +
  scale_y_continuous(breaks= pretty_breaks(n=6), limits = c(0,80)) +
  scale_color_manual(values = c("chocolate3","cadetblue3", "darkgreen" )) +
  scale_fill_manual(values = c("chocolate3", "cadetblue3", "darkgreen" )) +
  scale_shape_manual(values = c(21, 22, 24))
figure1

# @ 3 x 3.5 portrait
```


### Statistical analyses


#### Does crayfish sex affect responses?

```{r}
  ggplot(craycray, aes(x=interaction(Crayfish_Sex,Treatment_Period), y=Time_Active_percent)) +
  geom_point(aes(fill=Predator_Treatment, shape = Predator_Treatment), position = position_dodge(width=0.1), size = 3, color = "black") +
  labs(y = "Mean time active (±95% CI)", x="Treatment") + #creat labels... 
  theme_classic(base_size=12) +
  theme(legend.text = element_text(size = 12), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 12), axis.ticks = element_line(color = "black"), axis.title.x = element_blank()) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= c(.01,.1), legend.title = element_blank()) +
  scale_y_continuous(breaks= pretty_breaks(n=6), limits = c(0,80)) +
  scale_color_manual(values = c("chocolate3","cadetblue3", "darkgreen" )) +
  scale_fill_manual(values = c("chocolate3", "cadetblue3", "darkgreen" )) +
  scale_shape_manual(values = c(21, 22, 24))

# Here we only test activity in the "after" treatment period to assess this
sex_interaction_fixed <- glmmTMB(Time_Active_prop~Predator_Treatment*Crayfish_Sex + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)
sex_fixed <- glmmTMB(Time_Active_prop~Predator_Treatment + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)


#AIC comparison
print(aictab(list(sex_interaction_fixed, sex_fixed), c("sex_interaction_fixed", "sex_fixed")), LL = FALSE)

DHARMa::simulateResiduals(sex_interaction_fixed) |> plot()

Anova(sex_interaction_fixed)
```

#### 3-way interaction model
```{r}

# Fit the full model (including three-way interaction)
full_model <- glmmTMB(Time_Active_prop~Predator_Treatment*Treatment_Period*Population + (1|Crayfish_Treatment_ID), data=craycray)

# Define nested models

# Model 1: Without three-way interaction
model1 <- glmmTMB(Time_Active_prop ~ Predator_Treatment * Treatment_Period + Predator_Treatment * Population + Treatment_Period * Population + Predator_Treatment + Treatment_Period + Population + (1 | Crayfish_Treatment_ID), data=craycray)

# Model 2: Without Predator_Treatment:Treatment_Period interaction
model2 <- glmmTMB(Time_Active_prop ~ Predator_Treatment * Population + Treatment_Period * Population + Predator_Treatment + Treatment_Period + Population + (1 | Crayfish_Treatment_ID), data=craycray)

# Model 3: Without Predator_Treatment:Population interaction
model3 <- glmmTMB(Time_Active_prop ~ Predator_Treatment * Treatment_Period + Treatment_Period * Population + Predator_Treatment + Treatment_Period + Population + (1 | Crayfish_Treatment_ID), data=craycray)

# Model 4: Without Treatment_Period:Population interaction
model4 <- glmmTMB(Time_Active_prop ~ Predator_Treatment * Treatment_Period + Predator_Treatment * Population + Predator_Treatment + Treatment_Period + Population + (1 | Crayfish_Treatment_ID), data=craycray)

# Model 5: Without Predator_Treatment main effect
model5 <- glmmTMB(Time_Active_prop ~ Treatment_Period * Population + Treatment_Period + Population + (1 | Crayfish_Treatment_ID), data=craycray)

# Model 6: Without Treatment_Period main effect
model6 <- glmmTMB(Time_Active_prop ~ Predator_Treatment * Population + Predator_Treatment + Population + (1 | Crayfish_Treatment_ID), data=craycray)

# Model 7: Without Population main effect
model7 <- glmmTMB(Time_Active_prop ~ Predator_Treatment * Treatment_Period + Predator_Treatment + Treatment_Period + (1 | Crayfish_Treatment_ID), data=craycray)

# Model 8: Intercept only (null model)
model8 <- glmmTMB(Time_Active_prop ~ 1 + (1 | Crayfish_Treatment_ID), data=craycray)

# Fit all models and compare using anova
anova_results <- anova(model1, model2, model3, model4, model5, model6, model7, model8, full_model)

# Print the ANOVA table
anova_results


#AIC comparison
print(aictab(list(full_model, model1, model2, model3, model4, model5, model6, model7,  model8), c("full_model", "model1", "model2", "model3", "model4", "model5", "model6", "model7",  "model8")), LL = FALSE)

# Best model is # Model 7: Without Population main effect based on AIC comparisons

#Check for normality assumptions
#homogeneity of variance assumption
plot(fitted(model7), residuals(model7))
#leveneTest(data=sp, Proportion_escape_hatched~Age*Clutch_size)
#normality assumption
hist(residuals(model7))
qqnorm(residuals(model7))
qqline(residuals(model7))
shapiro.test(residuals(model7))

DHARMa::simulateResiduals(full_model) |> plot()


summary(model7)
Anova(model7)
```

Likelihood ratio tests
```{r}
#3-way interaction effect
anova(full_model, model1, test = "Chisq")
#Predator_Treatment:Treatment_Period interaction effect
anova(full_model, model2, test = "Chisq")
#Predator_Treatment:Population effect
anova(full_model, model3, test = "Chisq")
#Treatment_Period:Population effect
anova(full_model, model4, test = "Chisq")


#cue treatment effect
anova(full_model, model5, test = "Chisq")
#time period effect
anova(full_model, model6, test = "Chisq")
#population effect
anova(full_model, model7, test = "Chisq")
```


#### 2-way interaction model
```{r}
# Adjust proportions to avoid exact 0 or 1
craycray$Active_prop_adj <- pmin(pmax(craycray$Time_Active_prop, 0.0001), 0.9999)

# Then logit-transform
craycray$logit_Active_prop <- log(craycray$Active_prop_adj / (1 - craycray$Active_prop_adj))

full_model1 <- glmmTMB(Time_Active_prop ~ Predator_Treatment * Treatment_Period + (1 | Crayfish_Treatment_ID), dispformula = ~ Predator_Treatment, data=craycray)
full_logit_LMM_model <- glmmTMB(logit_Active_prop ~ Predator_Treatment * Treatment_Period + (1 | Crayfish_Treatment_ID), dispformula = ~ Predator_Treatment, data=craycray)

nointeraction <- glmmTMB(Time_Active_prop ~ Predator_Treatment + Treatment_Period + (1 | Crayfish_Treatment_ID), dispformula = ~ Predator_Treatment, data=craycray)
noperiod <- glmmTMB(Time_Active_prop ~ Predator_Treatment + (1 | Crayfish_Treatment_ID), dispformula = ~ Predator_Treatment, data=craycray)
notreatment <- glmmTMB(Time_Active_prop ~ Treatment_Period + (1 | Crayfish_Treatment_ID), dispformula = ~ Predator_Treatment, data=craycray)


DHARMa::simulateResiduals(full_model1) |> plot()
DHARMa::simulateResiduals(full_logit_LMM_model) |> plot()
```


Likelihood ratio tests
```{r}
#treatment x period interaction effect
anova(full_model1, nointeraction, test = "Chisq")
#time period effect
anova(full_model1, noperiod, test = "Chisq")
#treatment effect
anova(full_model1, notreatment, test = "Chisq")
```

##### Effect Sizes & Post hoc Comparisons on Model 7: Without Population main effect
```{r}
#Compute Estimated Marginal Means (EMMs) & effect sizes & CI
emms = emmeans(full_model1, pairwise~Treatment_Period|Predator_Treatment, type="response") #, adjust="Bonferoni"
emms
summary(emms, adjust="fdr", infer=TRUE)

emms = emmeans(full_model1, pairwise~Predator_Treatment|Treatment_Period, type="response") #, adjust="Bonferoni"
emms
summary(emms, adjust="fdr", infer=TRUE)
```






## PREDATOR AVOIDANCE


### Position score
```{r}
Position_score_summary = craycray_after_period_only %>% 
  group_by(Predator_Treatment) %>%
summarise(mean = mean(Position_Score, na.rm=T), SE= sd(Position_Score, na.rm=T)/sqrt(length(Position_Score)), SD = sd(Position_Score, na.rm=T), max = max(Position_Score, na.rm=T), min = min(Position_Score, na.rm=T), sample_size_unique=n_distinct(Position_Score, na.rm=T), sample_size=n())
Position_score_summary
```

```{r}
Position_score_plot = 
  ggplot(craycray_after_period_only, aes(x=Predator_Treatment, y=Position_Score)) +
  geom_boxplot(aes(color=Predator_Treatment), fill=NA, outlier.shape = NA) + 
  geom_point(aes(fill=Predator_Treatment), position = position_jitterdodge(.2), size = 2, color = "black", pch=21) +
  labs(y="Mean position score", x="Treatment") + #creat labels... 
  theme_classic(base_size=12) +
  theme(legend.text = element_text(size = 12), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 12), axis.ticks = element_line(color = "black"), axis.title.x = element_blank()) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= "none") +
  scale_y_continuous(breaks= pretty_breaks(n=7)) +
  scale_color_manual(values = c("chocolate3","cadetblue3", "darkgreen" )) +
  scale_fill_manual(values = c("chocolate3", "cadetblue3", "darkgreen" ))
Position_score_plot
```


#### STATS

#### Does crayfish sex affect responses?

```{r}
# Here we only test activity in the "after" treatment period to assess this
sex_interaction_fixed <- lmer(Position_Score~Predator_Treatment*Crayfish_Sex + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)
sex_fixed <- lmer(Position_Score~Predator_Treatment + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)

#AIC comparison
library(AICcmodavg)
print(aictab(list(sex_interaction_fixed, sex_fixed), c("sex_interaction_fixed", "sex_fixed")), LL = FALSE)

Anova(sex_interaction_fixed)
```



```{r}
library(lme4)
library(MuMIn)
library(car)

hist(craycray3$Position_Score)

#all models
mean_position_score_full = lmer(Position_Score~Predator_Treatment*Population + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)
mean_position_score_nointeraction=  lmer(Position_Score~Predator_Treatment+Population + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)
mean_position_score_notreatment = lmer(Position_Score~Population + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)
mean_position_score_notimeperiod = lmer(Position_Score~Predator_Treatment + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)


mean_position_score_full_sex = lmer(Position_Score~Predator_Treatment*Population + (1|Crayfish_Treatment_ID) + (1|Crayfish_Sex), data=craycray_after_period_only)
mean_position_score_full2 = lmer(log(Position_Score)~Predator_Treatment*Population + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)

#AIC comparison
library(AICcmodavg)
print(aictab(list(mean_position_score_full,  mean_position_score_nointeraction, mean_position_score_notreatment, mean_position_score_notimeperiod, mean_position_score_full_sex), c("mean_position_score_full", "mean_position_score_nointeraction", "mean_position_score_notreatment", "mean_position_score_notimeperiod", "mean_position_score_full_sex")), LL = FALSE)

#time_in_danger_zone_full = glmer(Time_in_Danger_Zone_prop~Predator_Treatment*Population + (1|Crayfish_Treatment_ID), family="binomial", weights=Total_Number_of_Frames, data=craycray3)



#Check for normality assumptions
#homogeneity of variance assumption
plot(fitted(mean_position_score_full), residuals(mean_position_score_full))
#leveneTest(data=sp, Proportion_escape_hatched~Age*Clutch_size)
#normality assumption
hist(residuals(mean_position_score_full))
qqnorm(residuals(mean_position_score_full))
qqline(residuals(mean_position_score_full))
shapiro.test(residuals(mean_position_score_full))

Anova(mean_position_score_full)
```

Likelihood ratio tests
```{r}
anova(mean_position_score_full, mean_position_score_nointeraction, test = "Chisq")
anova(mean_position_score_full, mean_position_score_notreatment, test = "Chisq")
anova(mean_position_score_full, mean_position_score_notimeperiod, test = "Chisq")
```

EFFECT OF INTERACTION:
* **X = 0.0068; p-value = 0.9966 **

EFFECT OF TREATMENT:
* **X = 11.139; p-value = 0.02504 **

EFFECT OF POPULATION:
* **X = 0.5096; p-value = 0.9168 **



##### Effect Sizes & Post hoc Comparisons
```{r}
emmeans(mean_position_score_full, pairwise ~ Predator_Treatment, adjust = "tukey")
```


## FEEDING RESPONSE

Summarize proportion data
```{r}
library(dplyr)

proportion_fed_summary = craycray_after_period_only %>%
  group_by(Predator_Treatment) %>%
  summarise(sample_size = n(), successes = sum(Feeding_Response), proportion_fed = successes/sample_size)
proportion_fed_summary

proportion_fed_summary2 = craycray_after_period_only %>%
  group_by(Predator_Treatment, Population) %>%
  summarise(sample_size = n(), successes = sum(Feeding_Response), proportion_fed = successes/sample_size)
proportion_fed_summary2

proportion_fed_summary3 = craycray_after_period_only %>%
  group_by(Predator_Treatment, Crayfish_Sex) %>%
  summarise(sample_size = n(), successes = sum(Feeding_Response), proportion_fed = successes/sample_size)
proportion_fed_summary3

#Control vs. Control
# Calculate the odds ratio and its 95% CI
epi.2by2(matrix(c(13, 0, 16, 1), ncol = 2), method = "cohort.count", conf.level = 0.95)

#Native
epi.2by2(matrix(c(3, 13, 3, 11), ncol = 2), method = "cohort.count", conf.level = 0.95)

#Invasive
epi.2by2(matrix(c(13, 3, 9, 2), ncol = 2), method = "cohort.count", conf.level = 0.95)
```

### Visualizations

Proportion fed
```{r}
library(ggplot2)
library(dplyr)

proportion_fed = ggplot(proportion_fed_summary, aes(y=proportion_fed, x=Predator_Treatment)) +
  geom_bar(stat="identity", aes(fill=Predator_Treatment), position=position_dodge(.9), width = .9, color="black") + # use barplots w/time of day filled w/color
  labs(y="Proportion that fed", x=element_blank()) + #creat labels... 
  theme_classic(base_size = 12) +
  theme(legend.position="none", axis.title.y = element_text(size = 12), axis.text = element_text(color="black", size = 12), axis.ticks = element_line(color = "black")) +
  scale_y_continuous(expand= c(0,0), breaks= pretty_breaks(n=6), limits = c(0,1.09)) +
  scale_color_manual(values = c("chocolate3","cadetblue3", "darkgreen" )) +
  scale_fill_manual(values = c("chocolate3", "cadetblue3", "darkgreen" )) +
  geom_text(aes(y= 0.025, label= successes), color= "black", vjust = 0, size = 4) 
proportion_fed
```

Latency to Feed Summary
```{r}
latency_to_feed_summary = craycray_after_period_only %>%
  group_by(Predator_Treatment) %>%
  summarise(mean = mean(Latency_to_Feed_min, na.rm=T), SE= sd(Latency_to_Feed_min, na.rm=T)/sqrt(length(Latency_to_Feed_min)), SD = sd(Latency_to_Feed_min, na.rm=T), max = max(Latency_to_Feed_min, na.rm=T), min = min(Latency_to_Feed_min, na.rm=T), sample_size_unique=n_distinct(Latency_to_Feed_min, na.rm=T), sample_size=n())
latency_to_feed_summary

latency_to_feed_summary2 = craycray_after_period_only %>%
  group_by(Population, Predator_Treatment) %>%
  summarise(mean = mean(Latency_to_Feed_min, na.rm=T), SE= sd(Latency_to_Feed_min, na.rm=T)/sqrt(length(Latency_to_Feed_min)), SD = sd(Latency_to_Feed_min, na.rm=T), max = max(Latency_to_Feed_min, na.rm=T), min = min(Latency_to_Feed_min, na.rm=T), sample_size_unique=n_distinct(Latency_to_Feed_min, na.rm=T), sample_size=n())
latency_to_feed_summary2

latency_to_feed_summary3 = craycray_after_period_only %>%
  group_by(Population) %>%
  summarise(mean = mean(Latency_to_Feed_min, na.rm=T), SE= sd(Latency_to_Feed_min, na.rm=T)/sqrt(length(Latency_to_Feed_min)), SD = sd(Latency_to_Feed_min, na.rm=T), max = max(Latency_to_Feed_min, na.rm=T), min = min(Latency_to_Feed_min, na.rm=T), sample_size_unique=n_distinct(Latency_to_Feed_min, na.rm=T), sample_size=n())
latency_to_feed_summary3
```


```{r}
library(ggplot2)
library(scales)
latency_to_feed = 
  ggplot(craycray_after_period_only, aes(x=Predator_Treatment, y=Latency_to_Feed_min)) +
  geom_boxplot(aes(color=Predator_Treatment), fill=NA, outlier.shape = NA) + 
  geom_point(aes(fill=Predator_Treatment), position = position_jitterdodge(.2), size = 2, color = "black", pch=21) +
  labs(y="Latency to feed (min)", x="Treatment") + #creat labels... 
  theme_classic(base_size=12) +
  theme(legend.text = element_text(size = 12), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 12), axis.ticks = element_line(color = "black"), axis.title.x = element_blank()) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= "none") +
  scale_y_continuous(breaks= pretty_breaks(n=6)) +
  scale_color_manual(values = c("chocolate3","cadetblue3", "darkgreen" )) +
  scale_fill_manual(values = c("chocolate3", "cadetblue3", "darkgreen" )) 
latency_to_feed
```


### Composite figure
```{r}
library(gridExtra)
library(cowplot)

#  @ portrait at 7.08661 x 8.26772 max for JEB
# 4x8.5 portrait looks good

Feeding_response = plot_grid(proportion_fed, latency_to_feed, labels = "AUTO", label_x = -.01, label_y = 1, scale =1, ncol=1, nrow = 2, label_size = 12, align = "hv")
Feeding_response

# @ 3 x 6
```

### STATS

#### Proportion fed
```{r}
#Native vs. Control
prop.test(x=c(6,29), n=c(30,30))
# Calculate the odds ratio and its 95% CI
epi.2by2(matrix(c(6, 24, 29, 1), ncol = 2), method = "cohort.count", conf.level = 0.95)

#Invasive vs. Control
prop.test(x=c(22,29), n=c(30,30))
# Calculate the odds ratio and its 95% CI
epi.2by2(matrix(c(22, 8, 29, 1), ncol = 2), method = "cohort.count", conf.level = 0.95)


#Native vs. Invasive
prop.test(x=c(6,22), n=c(30,30))
# Calculate the odds ratio and its 95% CI
epi.2by2(matrix(c(6, 24, 22, 8), ncol = 2), method = "cohort.count", conf.level = 0.95)

```


#### Latency to feed


#### Does crayfish sex affect responses?

```{r}
# Here we only test activity in the "after" treatment period to assess this
sex_interaction_fixed = lmer(log(Latency_to_Feed_min)~Predator_Treatment*Crayfish_Sex + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)
sex_fixed <- lmer(log(Latency_to_Feed_min)~Predator_Treatment + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)

#AIC comparison
library(AICcmodavg)
print(aictab(list(sex_interaction_fixed, sex_fixed), c("sex_interaction_fixed", "sex_fixed")), LL = FALSE)

Anova(sex_interaction_fixed)
```


GLMM
```{r}
library(lme4)
library(MuMIn)
library(car)

hist(craycray$Latency_to_Feed_min)
hist(log(craycray$Latency_to_Feed_min))


#all models
latency_to_feed_full = lmer(log(Latency_to_Feed_min)~Predator_Treatment*Population + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)
latency_to_feed_nointeraction = lmer(log(Latency_to_Feed_min)~Predator_Treatment+Population + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)
latency_to_feed_notreatment = lmer(log(Latency_to_Feed_min)~Population + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)
latency_to_feed_population = lmer(log(Latency_to_Feed_min)~Predator_Treatment + (1|Crayfish_Treatment_ID), data=craycray_after_period_only)

latency_to_feed_full_sex = lmer(log(Latency_to_Feed_min)~Predator_Treatment*Population + (1|Crayfish_Treatment_ID) + (1|Crayfish_Sex), data=craycray_after_period_only)

#AIC comparison
library(AICcmodavg)
print(aictab(list(latency_to_feed_full,  latency_to_feed_nointeraction, latency_to_feed_notreatment, latency_to_feed_population, latency_to_feed_full_sex), c("latency_to_feed_full",  "latency_to_feed_nointeraction", "latency_to_feed_notreatment", "latency_to_feed_population", "latency_to_feed_full_sex")), LL = FALSE)

#Check for normality assumptions
#homogeneity of variance assumption
plot(fitted(latency_to_feed_full), residuals(latency_to_feed_full))
#leveneTest(data=sp, Proportion_escape_hatched~Age*Clutch_size)
#normality assumption
hist(residuals(latency_to_feed_full))
qqnorm(residuals(latency_to_feed_full))
qqline(residuals(latency_to_feed_full))
shapiro.test(residuals(latency_to_feed_full))

Anova(latency_to_feed_full)
```

Likelihood ratio tests
```{r}
anova(latency_to_feed_full, latency_to_feed_nointeraction, test = "Chisq")
anova(latency_to_feed_full, latency_to_feed_notreatment, test = "Chisq")
anova(latency_to_feed_full, latency_to_feed_population, test = "Chisq")
```

EFFECT OF INTERACTION:
* **X = 1.0052; p-value = 0.6049 **

EFFECT OF TREATMENT:
* **X = 12.666; p-value = 0.01303**

EFFECT OF POPULATION:
* **X = 5.5689; p-value = 0.1346**


##### Effect Sizes & Post hoc Comparisons
```{r}
emmeans(latency_to_feed_full, pairwise ~ Predator_Treatment, adjust = "tukey")
```





******************************************************************************************************
******************************************************************************************************
******************************************************************************************************



# Water-borne Cue Trials

```{r}
wbc_df <- read_excel("Datasets for Invasive swamp eels do not elicit antipredator behavior in a native crayfish with collapsing populations.xlsx", 
    sheet = "Water-borne Cues Experiment", na = "NA")
wbc_df

wbc_df = wbc_df %>%
  mutate(Treatment_Period = factor(Treatment_Period, levels=c("Before", "After")))
```

## Sumary statistics
```{r}
waterborncues_summary1 = wbc_df %>% 
  group_by(Cue_Treatment, Treatment_Period) %>%
summarise(mean = mean(Time_Active_percent, na.rm=T), SE= sd(Time_Active_percent, na.rm=T)/sqrt(length(Time_Active_percent)), SD = sd(Time_Active_percent, na.rm=T), max = max(Time_Active_percent, na.rm=T), min = min(Time_Active_percent, na.rm=T), sample_size_unique=n_distinct(Time_Active_percent, na.rm=T), sample_size=n(), ci_lower = mean-qt(0.975, df = n() - 1) *SE, ci_upper = mean + qt(0.975, df = n() - 1) * SE)
waterborncues_summary1

waterborncues_summary2 = wbc_df %>% 
summarise(mean = mean(Crayfish_CL_mm, na.rm=T), SE= sd(Crayfish_CL_mm, na.rm=T)/sqrt(length(Crayfish_CL_mm)), SD = sd(Crayfish_CL_mm, na.rm=T), max = max(Crayfish_CL_mm, na.rm=T), min = min(Crayfish_CL_mm, na.rm=T), sample_size_unique=n_distinct(Crayfish_CL_mm, na.rm=T), sample_size=n(), ci_lower = mean-qt(0.975, df = n() - 1) *SE, ci_upper = mean + qt(0.975, df = n() - 1) * SE)
waterborncues_summary2
```


## Visualizations
```{r}
waterborncues_effect_sizes = 
  ggplot(waterborncues_summary1, aes(x=Treatment_Period, y=mean)) +
  geom_path(aes(color=Cue_Treatment, group = Cue_Treatment), position = position_dodge(width=0.1), size = 3) +
  geom_errorbar(aes(color=Cue_Treatment, ymin=mean-SE, ymax=mean+SE), position = position_dodge(width=0.1), width=0.5, size=2) +
  geom_point(aes(fill=Cue_Treatment, shape = Cue_Treatment), position = position_dodge(width=0.1), size = 4, color = "black") +
  #stat_summary(aes(color=Cue_Treatment, alpha=Population), fun=mean, geom='crossbar', position = position_jitterdodge(.1), size=1, linewidth=.5) +
  labs(y = "Mean time active (±95% CI)", x="Treatment") + #creat labels... 
  theme_classic(base_size=12) +
  theme(legend.text = element_text(size = 12), legend.title = element_blank(), legend.background = element_blank(), axis.text = element_text(color="black", size = 12), axis.ticks = element_line(color = "black"), axis.title.x = element_blank()) +
  theme(legend.title.align=0.5, legend.justification = c(0,.5), legend.position= c(.05,.1), legend.title = element_blank()) +
  scale_y_continuous(breaks= pretty_breaks(n=6), limits = c(0,80)) +
  scale_color_manual(values = c("darkgray", "chocolate3","cadetblue3", "darkgreen" )) +
  scale_fill_manual(values = c("darkgray", "chocolate3", "cadetblue3", "darkgreen" )) +
  scale_shape_manual(values = c(21, 22, 23, 24))
waterborncues_effect_sizes

# @3x3.5 portrait
```

### STATS

#### Does crayfish sex affect responses?

```{r}
wbc_after_only_df = wbc_df %>%
  filter(Treatment_Period== "After")

# Here we only test activity in the "after" treatment period to assess this
sex_interaction_fixed =  glmmTMB(Time_Active_prop~Cue_Treatment*Crayfish_Sex + (1|Crayfish_Treatment_ID), data=wbc_after_only_df)
sex_fixed <- glmmTMB(Time_Active_prop~Cue_Treatment + (1|Crayfish_Treatment_ID), data=wbc_after_only_df)

#AIC comparison
print(aictab(list(sex_interaction_fixed, sex_fixed), c("sex_interaction_fixed", "sex_fixed")), LL = FALSE)

Anova(sex_interaction_fixed)
```

```{r}
wbc_df_nocontrol = wbc_df %>%
  filter(Cue_Treatment!= "Control")


activity_water_full_model <- glmmTMB(Time_Active_prop~Cue_Treatment*Treatment_Period + (1|Crayfish_Treatment_ID), data=wbc_df)

activity_water_nointeraction = glmmTMB(Time_Active_prop~Cue_Treatment+Treatment_Period + (1|Crayfish_Treatment_ID), data=wbc_df)
activity_water_notreatment = glmmTMB(Time_Active_prop~Treatment_Period + (1|Crayfish_Treatment_ID), data=wbc_df)
activity_water_period = glmmTMB(Time_Active_prop~Cue_Treatment + (1|Crayfish_Treatment_ID), data=wbc_df)


#AIC comparison
print(aictab(list(activity_water_full_model,  activity_water_nointeraction, activity_water_notreatment, activity_water_period), c("activity_water_full_model",  "activity_water_nointeraction", "activity_water_notreatment", "activity_water_period")), LL = FALSE)

#Check for normality assumptions
#homogeneity of variance assumption
plot(fitted(activity_water_full_model), residuals(activity_water_full_model))
#leveneTest(data=sp, Proportion_escape_hatched~Age*Clutch_size)
#normality assumption
hist(residuals(activity_water_full_model))
qqnorm(residuals(activity_water_full_model))
qqline(residuals(activity_water_full_model))
shapiro.test(residuals(activity_water_full_model))

DHARMa::simulateResiduals(activity_water_full_model) |> plot()

Anova(activity_water_full_model)
```


Likelihood ratio tests
```{r}
anova(activity_water_full_model, activity_water_nointeraction, test = "Chisq")
anova(activity_water_full_model, activity_water_notreatment, test = "Chisq")
anova(activity_water_full_model, activity_water_period, test = "Chisq")
```

